---
title: "Traditional agroforestry systems in Timor-Leste can store large amounts of carbon in both soil and biomass"
author: "Marguerite Cogné, Régis Peltier, Vincent Freycon, Alexis Toumazeau, Camille Piponiot"
date: "2023-05-10"
output: 
  bookdown::word_document2:
  fig_caption: yes
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results = "hide", tinytex.clean = FALSE)
packages_needed = c("ggplot2", "dataverse", "readxl", "data.table")

packages_to_install = packages_needed[!( packages_needed %in% rownames(installed.packages()))]

if (length(packages_to_install) > 0)
  install.packages(packages_to_install)

lapply(packages_needed, require, character.only = TRUE)
```

# Introduction

Agricultural expansion has been the main driver of tropical deforestation over the past half century [@Pendrill2022;@Gibbs2010;@Rudel2009], resulting in large greenhouse gas emissions [approximately 2.6 gigatonnes CO_2_e/yr over 2010-2014; @Pendrill2019]. The demand for tropical agricultural land is expected to continue to increase in the tropics as the population grows [@UN2022]. In addition, tropical areas are expected to be disproportionately affected by deforestation and climate change, putting their agriculture at risk [@Lawrence2015]. There is therefore an urgent need to promote tropical agricultural landscapes that are both more resilient to climate change, while maintaining high productivity and ecological value [@Harvey2014]. In particular, increasing the carbon stored in the soil and vegetation of agricultural landscapes could be an effective way to mitigate climate change while increasing the resilience of the food production system [@Lal2004].

Agroforestry has attracted attention in recent decades as a set of effective agricultural practices that could contribute to climate change mitigation and adaptation [@Verchot2007]. Agroforestry is a collective name for land use systems and technologies where woody perennials (trees, shrubs, palms, bamboo, etc.) are deliberately used on the same land management units as agricultural crops and/or animals, in some form of spatial arrangement or temporal sequence [@Lundgren1982]. Agroforestry allows for diversification of agricultural production, making agricultural systems more resilient [@Duffy2021]; it also increases the carbon sequestration in the vegetation but also in the soil [@DeStefano2018].  


Carbon stocks in AFS [VF]

Importance of the ecological/social context in the success of agroforestry techniques - existence of traditional agroforestry techniques that could be more adapted to local conditions, [MC]
Agroforestry needs to be adapted to the context [(Coe et al., 2014, Cur Opinion in Env Sust) @Coe2014]

Existence of traditional agroforestry systems [examples from Marguerite’s intro]

Context in Timor Leste: current knowledge on their agroforestry practices + knowledge gap [MC]

Local typology of production systems and why it is not adapted to traditional agroforestry

No prior description of traditional agroforestry systems in TL

In Timor Leste, a typology of production systems taking into account different agroecological zoning was carried out in 2017 based on the results of the 2010 national census (R. Williams et al, 2018). However, this typology does not highlight the agroforestry component of Timorese agriculture by focusing on other differentiating criteria such as livestock, subsistence crops (rice) or the island's historical cash crops (coconut, coffee).

Need to better understand how these systems work and quantify their carbon storage to include them in climate change mitigation policies 
Research questions  [VF]

What are the characteristics of traditional agroforestry systems in Timor Leste? 

How much carbon do they store in both their biomass and in the soil

Are there measurable differences between AFS?

Hypotheses [VF]

Proposition: lien entre typologie et stocks de carbone

# Materials and methods

```{r download-data}
## download data from dataverse - once published
dataverse_url <- "https://dataverse.cirad.fr/privateurl.xhtml?token=f8eabb2a-3635-499b-aee5-80fd2b7495a0"
dataverse_doi <- "10.18167/DVN1/QCXWIY"

# for now : download manually from temporary link
# utils::unzip("dataverse_files.zip")
``` 

```{r agb-calculation}
# get agb data
source("script-agb.R")
```

```{r open-soil-data}
# open organic carbon data
# bulk density in Gariuai according to SoilGrids = 125 cg/cm3
data_soil <- readxl::read_excel("data/RESUME_DATA_SOIL_SHORT_TL.xls", sheet = "RESUME-LAB ")

# add bulk density and coarse fraction
data_soil2 <- readxl::read_excel("data/RAW_DATA_SOIL_LARGE_TL.xls", sheet = "SAMPLING_MEASURE")

# bulk density from direct measurements
data_soil2$BD <- data_soil2$Sample_DRY/data_soil2$V_cylinder

# coarse fraction
data_soil2$CF <- data_soil2$V_rocks_tot/data_soil2$V_cylinder

# change format of depth to correspond to other soil table
data_soil2$SAMPLE_DEPTH <- factor(data_soil2$DEPTH_SAMPLE_C, levels = c("2_7", "12_17", "22_27"))
levels(data_soil2$SAMPLE_DEPTH) <- c("0-10", "10-20", "20-30")

data_soil <- merge(data_soil, data_soil2, by = c("N_PLOT", "SAMPLE_DEPTH"), all=TRUE)
data.table::setDT(data_soil)
```

```{r test-bd-pedotransfer}
# bulk density from pedotransfer function: Minasny and Hartemink, 2011

data_soil$depth_mean <- unlist(lapply(strsplit(data_soil$SAMPLE_DEPTH, "-"), function(x) mean(as.numeric(x))))
data_soil$depth_max <- unlist(lapply(strsplit(data_soil$SAMPLE_DEPTH, "-"), function(x) as.numeric(x[2])))

data_soil$sand <- as.numeric(data_soil$`FINE SAND`) + as.numeric(data_soil$`COARSE SAND`)/100

data_soil$BDmin <- 
  0.935 + 
  0.049 * log(data_soil$depth_mean) + 
  0.0055 * data_soil$sand + 
  0.000065 * (data_soil$sand - 38.96)^2

data_soil$BDptf <- 
  100/(
    data_soil$OM/0.224 + 
      (100 - data_soil$OM)/data_soil$BDmin
  )

# g1 = ggplot(data_soil, aes(x = BD, y = BDptf, color = ID_TYPE)) + 
#   geom_abline(slope = 1, intercept = 0, lty=2) + 
#   geom_point() + 
#   scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
#   lims(x = c(0.8, 1.6), y = c(0.8, 1.6)) +
#   facet_wrap(~ID_VILLAGE)
# #+
#   # theme(legend.position = "none")
# g2 = ggplot(data_soil, aes(x = BD, y = BDmin, color = SAMPLE_DEPTH)) +
#   geom_abline(slope = 1, intercept = 0, lty=2) + 
#   geom_point() + 
#   scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
#   lims(x = c(0.8, 1.6), y = c(0.8, 1.6)) +
#   theme(legend.position = c(0.2, 0.8))
# 
# ggpubr::ggarrange(g1, g2) 
# ggsave("figures/pedotransfer-test.png", height = 5, width = 10)
```

```{r add-missing-values}
## add intermediate soil carbon in 10-20 cm depth (for now: no value)
data_soil[, Corg := mean(`Corg\n`, na.rm = TRUE), .(N_PLOT)]
data_soil[!is.na(`Corg\n`), Corg := `Corg\n`]

# replace one missing value of coarse fraction with plot level mean
data_soil[, CF_site := mean(CF, na.rm = TRUE), .(N_PLOT)]
data_soil[is.na(CF), CF := CF_site]
```

```{r total-soc}
# calculate total soil carbon stock per depth
data_soil[, c("h0", "h1") := lapply(data.table::tstrsplit(SAMPLE_DEPTH, "-"), as.numeric)]
data_soil[, Cstock :=          ## in Mg/ha
            (h1 - h0) * 1e8 *  ## equivalent volume of a 1 ha soil sample, in cm3/ha
            BD *               ## bulk density 
            Corg/100 *         ## carbon organic content in soil
            (1-CF) *           ## remove coarse fraction
            1e-6               ## convert from g/ha to Mg/ha
]

data_soil_C <-
  data_soil[, .(
    meanSC = sum(Cstock),
    SC10 = Cstock[SAMPLE_DEPTH == "0-10"],
    clay = mean(as.numeric(CLAY), na.rm = TRUE)
  ), .(N_PLOT)]

setkey(data_soil_C, "N_PLOT")
```

```{r soc-error-propagation}
## error propagation using a Monte Carlo method -------------
data_soil_unc = data_soil[, .(BD = rnorm(1000, BD, BD*0.1), 
                              Corg = rnorm(1000, Corg, Corg*0.12), 
                              CF = rnorm(1000, CF, CF*0.15), 
                              iter = 1:1000), 
                          .(N_PLOT, SAMPLE_DEPTH, h1, h0)]

data_soil_unc = data_soil_unc[, .(
  Cstock = sum(
    (h1 - h0) * 1e8 *  
      BD * Corg/100 * (1-CF) * 1e-6  
  ), 
  Cstock10 = (
    (h1 - h0) * 1e8 *  
      BD * Corg/100 * (1-CF) * 1e-6  
  )[SAMPLE_DEPTH == "0-10"]),
  .(N_PLOT, iter)
]

data_soil_unc = data_soil_unc[, .(
  lwrSC = quantile(Cstock, 0.025), 
  uprSC = quantile(Cstock, 0.975),
  lwrSC10 = quantile(Cstock10, 0.025), 
  uprSC10 = quantile(Cstock10, 0.975)), 
  .(N_PLOT)]
```

```{r combine-agb-soc-data}
# combine AGB and soil C
dataC <- merge(data_agb, data_soil_C, by = "N_PLOT", all=TRUE)

# add soil carbon uncertainty
dataC <- merge(dataC, data_soil_unc, all=TRUE)

dataC[N_PLOT==6, `:=`(ID_VILLAGE = "GD", ID_AF = "CF")]

dataC[, village := substr(ID_VILLAGE, 1, 1)]
dataC$ID_AF <- factor(dataC$ID_AF, levels = c("CF", "SP", "YA", "HG", "FG"))
```

# Results

```{r tab-mean, results = "show"}
# average values per village and SAF type
dataC_mean <- dataC[, .(AGC = paste0(round(mean(meanAGC[!is.na(meanAGC)]), 2), " +/- ", round(sd(meanAGC[!is.na(meanAGC)])/sqrt(length(meanAGC[!is.na(meanAGC)])), 2)), 
                        SC = paste0(round(mean(meanSC[!is.na(meanSC)]), 2), " +/- ", round(sd(meanSC[!is.na(meanSC)])/sqrt(length(meanSC[!is.na(meanSC)])), 2))),
                    .(ID_AF)]
dataC_mean$ID_AF <- factor(dataC_mean$ID_AF, levels = c("CF", "SP", "YA", "HG", "FG"))
dataC_mean = dataC_mean[order(dataC_mean$ID_AF),]
knitr::kable(dataC_mean, col.names = c("Agroforestry system", "Aboveground carbon (Mg/ha)", "Soil organic carbon (Mg/ha)"))
write.csv(dataC_mean, file = "tables/C_mean_SAF.csv", row.names = FALSE)
```

```{r tab-anova, results = "show"}
# 2-way anova 
dataC[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
anov1 = anova(lm(meanAGC~ID_SUCO+ID_AF, data = dataC))
anov2 = anova(lm(meanSC~ID_SUCO+ID_AF, data = dataC))
tab_anov = data.frame(response = rep(c("AGC", "SOC"), each = 2), 
                      ind_var = rep(c("SUCO", "AF"), 2), 
                      p = round(c(anov1$`Pr(>F)`[1:2], anov2$`Pr(>F)`[1:2]), 3)
)
```

```{r fig-1-map, fig.height =6, fig.width = 7}

## 2. Load data from excel file ####
data_coord <- read_xlsx("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = "GPS")
data.table::setDT(data_coord)

data_coord[, N_PLOT := as.numeric(data.table::tstrsplit(N_Q_ID, "_")[[1]])]

data_coord <- data_coord[, .(long = mean(as.numeric(LONG)), lat = mean(as.numeric(LAT))), .(N_PLOT)]

data_plot <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "PLOT_GENERAL_INFO")

data_plot <- merge(data_coord, data_plot[, c("N_PLOT", "ID_VILLAGE", "ID_AF")])

data_plot[, village := substr(ID_VILLAGE, 1, 1)]
data_plot[village == "S", village := "Osso Luga"]
data_plot[village == "G", village := "Gariuai"]

data_box = data_plot[, .(xc = mean(long), yc = mean(lat)), .(village)]
data_box[village == "Gariuai", `:=`(w = 0.1, xc = xc + 0.001)]
data_box[village == "Osso Luga", w := 0.05]
data_box[, `:=`(x1 = xc - w/2, x2 = xc + w/2, y1 = yc - w/2, y2 = yc + w/2)]

## create Timor map ####

bbox <- c(124.5, -9.2, 127.5,  -8.3)
# basemap <- ggmap::get_stamenmap(bbox, maptype = "terrain", zoom = 10)

# geodata::elevation_3s(lon= c(data_box[village == "Osso Luga"]$x1), 
#                       lat= c(data_box[village == "Osso Luga"]$y1), 
#                       path = "data")
timor_map <- map_data("world", region = "Timor-Leste")

g1 = ggplot(timor_map) + 
  annotate(geom = "text", label = "a", x = -Inf, y = Inf, 
           hjust = 0, vjust = 1, fontface = "bold", size = 6) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill="lightgray", colour = "grey") +
  geom_tile(data = data_box, aes(x = xc, y = yc, width = w, height = w), fill = NA, color = 1, linewidth = 0.5) +
  ggrepel::geom_text_repel(data = data_box, aes(x = xc, y = yc, label = village)) + 
  labs(x = "", y = "") +
  theme(legend.position = "top", axis.title = element_blank(), 
        axis.ticks = element_blank(), axis.text = element_blank(), 
        panel.background = element_blank())+
  coord_sf(crs = 'WGS84')


# get elevation information
elev = as.data.frame(terra::rast("data/srtm_62_14.tif"), xy = TRUE)
elevS <- subset(elev, x > data_box[village == "Osso Luga"]$x1 & 
                  x < data_box[village == "Osso Luga"]$x2 & 
                  y > data_box[village == "Osso Luga"]$y1 & 
                  y < data_box[village == "Osso Luga"]$y2)
elevS$village = "Osso Luga"
elevG <- subset(elev, x > data_box[village == "Gariuai"]$x1 & 
                  x < data_box[village == "Gariuai"]$x2 & 
                  y > data_box[village == "Gariuai"]$y1 & 
                  y < data_box[village == "Gariuai"]$y2)
elevG$village = "Gariuai"

elevAll = rbind(elevS, elevG)

## Map of villages and plots with elevation
g2 = 
  ggplot(data = elevAll) +
  geom_raster(aes(x = x, y = y, fill=srtm_62_14)) +
  geom_point(data = data_plot, 
             aes(x = long, y = lat, color = ID_AF), size = 2) + 
  geom_point(data = data_plot, 
             aes(x = long, y = lat), shape = 1, size = 2) + 
  # coord_equal() +
  viridis::scale_fill_viridis(option = "D",
                              limits = range(c(elevS$srtm_62_14, elevG$srtm_62_14))
  ) +
  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
  labs(x = NULL, y = NULL, color = NULL, fill="Elevation (m)") +
  annotate(geom = "text", label = "b", x = -Inf, y = Inf, 
           hjust = -0.5, vjust = 1, fontface = "bold", size = 6) + 
  annotate(geom = "text", label = "c", x = Inf, y = -Inf, 
           hjust = 12, vjust = -9.8, fontface = "bold", size = 6) + 
  guides(color = FALSE, fill=guide_colorbar(title.position = "top")) +
  theme(legend.position = c(0.8, 0.85), legend.direction="horizontal", 
        axis.title = element_blank(), 
        axis.ticks = element_blank(), axis.text = element_blank(), 
        plot.title = element_text(hjust = 0.5), 
        panel.background = element_blank()) +
  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
  ggspatial::annotation_scale() +
  coord_sf(crs = 'WGS84')

# get AFS legend
g3 = ggplot(data_plot) +
  geom_point(aes(x = long, y = lat, color = ID_AF)) + 
  labs(color = "AFS") +
  guides(color=guide_legend(ncol=2)) +
  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
  theme_classic()
l1 = ggpubr::get_legend(g3)

ggpubr::ggarrange(ggpubr::ggarrange(g1, l1), g2, ncol = 1, heights = c(1,2))


ggsave("figures/mapPlotsSRTM.png", height =4.5, width = 6)


# extract elevation and rainfall for the 2 villages
data_plot$prec <- terra::extract(
  terra::rast("D:/maps/CHELSA/CHELSA_bio12_1981-2010_V.2.1.tif"), 
  cbind(data_plot$long, data_plot$lat)
)
data_plot$elev <- terra::extract(
  terra::rast("data/srtm_62_14.tif"), 
  cbind(data_plot$long, data_plot$lat)
)
data_plot[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
data_plot[, .(prec = mean(prec), range_prec = paste(range(prec), collapse = " - ")), .(ID_SUCO)]
data_plot[, .(elev = mean(elev), range_elev = paste(range(elev), collapse = " - ")), .(ID_SUCO)]
```


```{r anova}
## identify outliers
# library(magrittr)
library(rstatix)
dataC %>% 
  group_by(ID_AF) %>%
  identify_outliers(meanAGC)

dataC %>% 
  group_by(ID_AF) %>%
  identify_outliers(meanSC)

# remove extreme outlier and test normality
dataC[N_PLOT != 11] %>%
  group_by(ID_AF) %>%
  shapiro_test(meanAGC)

dataC %>%
  group_by(ID_AF) %>%
  shapiro_test(meanSC)

# non parametric kruskal wallis test 
kruskalAGC = kruskal.test(meanAGC ~ ID_AF,  data = dataC)
kruskal.test(meanSC ~ ID_AF,  data = dataC)
kruskal.test(meanAGC ~ village,  data = dataC)
kruskal.test(meanSC ~ village,  data = dataC)

# pairwise group comparison with Tukey HSD method
# https://arc.lib.montana.edu/book/statistics-with-r-textbook/item/59
library(multcomp)
aovAGC = aov(meanAGC ~ ID_AF, data = dataC[N_PLOT != 11])
tuk <- glht(aovAGC, linfct = mcp(ID_AF = "Tukey"))
summary(tuk)          # standard display
tuk.cld <- cld(tuk)   # letter-based display

plot(tuk.cld)


# soc
aovSOC = aov(meanSC ~ ID_AF, data = dataC)
tuk <- glht(aovSOC, linfct = mcp(ID_AF = "Tukey"))
summary(tuk)          # standard display
tuk.cld <- cld(tuk)   # letter-based display

opar <- par(mai=c(1,1,1.5,1))
plot(tuk.cld)

## non parametric test
#updated translation of help page implementation of Nemenyi-Damico-Wolfe-Dunn test
library(coin)
is.factor(dataC$ID_AF)
NDWD <-
  independence_test(meanAGC ~ ID_AF, data = dataC)
# , distribution = approximate(B = 10000),
# ytrafo = function(data) trafo(data, numeric_trafo = rank_trafo),
# xtrafo = mcp_trafo(iv = "Tukey"))

### global p-value
print(pvalue(NDWD))

### sites (I = II) != (III = IV) at alpha = 0.01 (page 244)
print(pvalue(NDWD, method = "single-step"))

#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-

library(tidyverse)
library(ggpubr)
library(rstatix)

# Data preparation
# 
# Here, we’ll use the built-in R data set named PlantGrowth. It contains the weight of plants obtained under a control and two different treatment conditions.

# Re-order the group levels:

# dataC <- dataC %>%
#   reorder_levels(group, order = c("ctrl", "trt1", "trt2"))

dataC %>% 
  group_by(ID_AF) %>%
  get_summary_stats(meanAGC, type = "common")


# Create a box plot of weight by group:


# Question: We want to know if there is any significant difference between the
# average weights of plants in the 3 experimental conditions.

# We’ll use the pipe-friendly kruskal_test() function [rstatix package], a
# wrapper around the R base function kruskal.test().

res.kruskal <- dataC %>% kruskal_test(meanAGC ~ ID_AF)
res.kruskal

# Effect size
#
# The eta squared, based on the H-statistic, can be used as the measure of the
# Kruskal-Wallis test effect size. It is calculated as follow : eta2[H] = (H - k
# + 1)/(n - k); where H is the value obtained in the Kruskal-Wallis test; k is
# the number of groups; n is the total number of observations (M. T. Tomczak and
# Tomczak 2014).
#
# The eta-squared estimate assumes values from 0 to 1 and multiplied by 100
# indicates the percentage of variance in the dependent variable explained by
# the independent variable.
#
# The interpretation values commonly in published literature are: 0.01- < 0.06
# (small effect), 0.06 - < 0.14 (moderate effect) and >= 0.14 (large effect).

dataC %>% kruskal_effsize(meanAGC ~ ID_AF)

# Multiple pairwise-comparisons
# 
# From the output of the Kruskal-Wallis test, we know that there is a significant difference between groups, but we don’t know which pairs of groups are different.
# 
# A significant Kruskal-Wallis test is generally followed up by Dunn’s test to identify which groups are different. It’s also possible to use the Wilcoxon’s test to calculate pairwise comparisons between group levels with corrections for multiple testing.


# Pairwise comparisons
pwc <- dataC %>% 
  dunn_test(meanAGC ~ ID_AF, p.adjust.method = "bonferroni") 
pwc[pwc$p.adj.signif=="*",]
```

```{r fig-2-barplot, fig.height = 5, fig.width = 6}
## melt data.table
dataC_all <- melt(dataC, id.vars = c("N_PLOT", "ID_VILLAGE", "ID_AF", "village"), 
                  measure.vars = c("meanAGC", "meanBGC", "meanSC"))

dataC_all <- dataC_all[, .(value = mean(value, na.rm = TRUE)), .(village, ID_AF, variable)]

## reorder factors 

dataC_all$ID_AF <- factor(dataC_all$ID_AF, levels = c("CF", "SP", "YA", "HG", "FG"))

# uncertainty fig 2: bootstrap by MC iteration?

dataC_all[, variable := factor(variable, levels = c("meanAGC", "meanBGC", "meanSC"))]
levels(dataC_all$variable) <- c("AGC", "BGC", "SOC")
dataC_all[, village := factor(village)]
levels(dataC_all$village) <- c("Gariuai", "Osso Luga")

ggplot(dataC_all, aes(x = ID_AF, y = value, fill = variable)) +
  geom_bar(position="stack", stat="identity") + 
  facet_wrap(~village, nrow = 1, strip.position = "bottom") +
  scale_fill_manual(values = c("forestgreen", "darkgoldenrod1", "coral4")) +
  theme_classic() +
  theme(strip.placement = "outside") +
  labs(x = "", y = "Carbon stocks (Mg C/ha)", fill = "Carbon\npool") +
  scale_y_continuous(expand = c(0, 0))
ggsave("figures/fig-carbon-stocks-stack.png", height = 5, width = 6)
```

```{r fig-3-reg, fig.height = 4, fig.width = 8}
## mixed effect linear regression
lmeReg <- lmerTest::lmer(meanSC ~ meanAGC + (1|village), data = dataC)
# Nakagawa’s R2 for mixed models
R2 = performance::r2_nakagawa(lmeReg)
summary(lmeReg)

# same for topsoil layer (0-10 cm)
lmeReg10 <- lmerTest::lmer(SC10 ~ meanAGC + (1|village), data = dataC)
# Nakagawa’s R2 for mixed models
R210 = performance::r2_nakagawa(lmeReg10)
summary(lmeReg10)

dfreg = 
  data.frame(
    depth = c("Soil depth: 0-30 cm", "Soil depth: 0-10 cm"), 
    intercept = c(summary(lmeReg)$coefficients[1,1], summary(lmeReg10)$coefficients[1,1]), 
    slope = c(summary(lmeReg)$coefficients[2,1], summary(lmeReg10)$coefficients[2,1]), 
    R2_cond = c(R2$R2_conditional, R210$R2_conditional), 
    R2_marg = c(R2$R2_marginal, R210$R2_marginal)
  )


# change village names
dataC[, village := factor(village)]
levels(dataC$village) <- c("Gariuai", "Osso Luga")

# stack data for figure

dataC_30 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "meanSC", "lwrSC", "uprSC")]
dataC_30$depth = "Soil depth: 0-30 cm"
dataC_10 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "SC10", "lwrSC10", "uprSC10")]
dataC_10$depth = "Soil depth: 0-10 cm"
colnames(dataC_10) = colnames(dataC_30)
dataC_fig = rbind(dataC_10, dataC_30)

## figure
ggplot(dataC_fig, aes(color = ID_AF)) +
  geom_point(aes(x = meanAGC, y = meanSC, shape = village), size = 2) + 
  geom_text(data = dfreg, aes(label = paste0("R^{2}*' (marg.) = '*", round(R2_marg, 3))),
            x = Inf, y = Inf, hjust = 1, vjust = 1.2, parse = TRUE, color = 1) +
  geom_text(data = dfreg, aes(label = paste0("R^{2}*' (cond.) = '*", round(R2_cond, 3))),
            x = Inf, y = Inf, hjust = 1, vjust = 2.2, parse = TRUE, color = 1) +
  geom_linerange(aes(x = meanAGC, y = meanSC, ymin = lwrSC, ymax = uprSC)) +
  geom_errorbarh(aes(y = meanSC, xmin = lwrAGC, xmax = uprAGC)) +
  geom_abline(data = dfreg, aes(intercept = intercept, slope = slope), lty = 2) +
  labs(color = "AF system", x = "Aboveground C (Mg C/ha)", y = "Soil C (Mg C/ha)") + 
  expand_limits(x = 0, y = 0) +
  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(~depth) + 
  theme_classic() 

ggsave("figures/fig-scatterplot-carbon.png", height = 4, width = 8)
```

```{r fig-3-reg-10, fig.height = 5, fig.width = 7}
## mixed effect linear regression
lmeReg10 <- lmerTest::lmer(SC10 ~ meanAGC + (1|village), data = dataC)
# Nakagawa’s R2 for mixed models
R2 = performance::r2_nakagawa(lmeReg)

## figure
ggplot(dataC, aes(color = ID_AF)) +
  geom_point(aes(x = meanAGC, y = SC10, shape = village), size = 2) + 
  annotate(geom= "text", x = Inf, y = Inf, hjust = 1, vjust = 1, parse = TRUE,
           label = paste0("R^{2}*' (marg.) = '*", round(R2$R2_marginal, 3))) +
  annotate(geom= "text", x = Inf, y = Inf, hjust = 1, vjust = 2, parse = TRUE,
           label = paste0("R^{2}*' (cond.) = '*", round(R2$R2_conditional, 3))) +
  # geom_linerange(aes(x = meanAGC, y = SC10, ymin = lwrSC, ymax = uprSC)) +
  geom_errorbarh(aes(y = SC10, xmin = lwrAGC, xmax = uprAGC)) +
  geom_abline(intercept = summary(lmeReg)$coefficients[1,1], slope = summary(lmeReg)$coefficients[2,1], lty = 2) +
  labs(color = "AF system", x = "Aboveground C (Mg C/ha)", y = "Soil C (Mg C/ha)") + 
  expand_limits(x = 0, y = 0) +
  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
  # scale_linetype_manual(values=c("twodash", "dotted"))+
  theme_classic() 
ggsave("figures/fig-scatterplot-carbon.png", height = 5, width = 7)
```

# Supplementary material 

```{r fig-distr-sol, fig.height = 5, fig.width = 15}
## diagnostic données sol pour marguerite
g1 <- ggplot(data_soil, aes(x = CF*100, col = SAMPLE_DEPTH)) + 
  geom_histogram(bins = 30) +
  geom_text(data = subset(data_soil, CF > 0.3), angle = -60, 
            aes(x = CF*100-5, y = 15, 
                label = paste("Village:", ID_VILLAGE, "/ SAF:", ID_TYPE, "/ Plot:", N_PLOT))) +
  labs(x = "Elements grossiers (% du volume)", 
       y = "Nombre d'échantillons (site x profondeur)", 
       col = "Profondeur (cm)") + 
  scale_y_continuous(expand = expand_scale(0, 0)) +
  theme_classic()  +
  theme(legend.position = c(0.8, 0.9))

g2 <- ggplot(data_soil, aes(x = Corg, col = SAMPLE_DEPTH)) + 
  geom_histogram(bins = 20) +
  labs(x = "Teneur en carbone (%)", 
       y = "Nombre d'échantillons (site x profondeur)", 
       col = "Profondeur (cm)") + 
  scale_y_continuous(expand = expand_scale(0, 0)) +
  theme_classic() +
  theme(legend.position = c(0.8, 0.9))

g3 <- ggplot(data_soil, aes(x = BD, col = SAMPLE_DEPTH)) + 
  geom_histogram(bins = 20) +
  labs(x = "Densité apparente", 
       y = "Nombre d'échantillons (site x profondeur)", 
       col = "Profondeur (cm)") + 
  scale_y_continuous(expand = expand_scale(0, 0)) +
  theme_classic()+
  theme(legend.position = c(0.8, 0.9))

ggpubr::ggarrange(g1, g2, g3, labels = "auto", nrow = 1)
ggsave("figures/fig-distr-sol.pdf", height = 5, width = 15)
```

```{r fig-clay-corg, fig.height = 4, fig.width = 6}
ggplot(data_soil, aes(x = as.numeric(CLAY), y = Corg, col = SAMPLE_DEPTH)) +
  geom_point() +
  labs(col = "Profondeur (cm)", x = "Argile (%)", y = "Teneur en carbone (%)") + 
  geom_smooth(method = "lm") +
  theme_classic()
ggsave("figures/fig-clay-corg.pdf", height = 4, width = 6)
```

```{r test-ac-soc-by-depth}
data_soil_agc <- merge(data_soil[, c("N_PLOT", "SAMPLE_DEPTH", "Cstock")], 
                       dataC[, c("N_PLOT","ID_AF", "ID_VILLAGE",  "meanAGC")], by = "N_PLOT")

ggplot(data_soil_agc, aes(x = meanAGC, y = Cstock, col = SAMPLE_DEPTH)) +
  geom_point() +
  geom_smooth(method = "lm")

lm(data = data_soil_agc, Cstock ~ meanAGC * SAMPLE_DEPTH) |> summary()
lm(data = data_soil_agc[SAMPLE_DEPTH == "0-10"], Cstock ~ meanAGC) |> summary()
lm(data = data_soil_agc[SAMPLE_DEPTH == "10-20"], Cstock ~ meanAGC ) |> summary()
lm(data = data_soil_agc[SAMPLE_DEPTH == "20-30"], Cstock ~ meanAGC ) |> summary()
```

```{r compare-lit-values}
litval <- read_excel("data/literature-values.xlsx")
colnames(litval) <- c("system", "pool", "value", "location", "ref")


litval$main_crop <- NA
litval$main_crop[grepl("palm", litval$system)] <- "oil palm"
litval$main_crop[grepl("offee", litval$system)] <- "coffee"
unique(litval$system[is.na(litval$main_crop)])

litval$main_per <- NA
litval$main_per[grepl("nset", litval$system)] <- "enset"
litval$main_per[grepl("Inga", litval$system)] <- "inga"
litval$main_per[grepl("Eucalyptus", litval$system)] <- "eucalyptus"
litval$main_per[grepl("Pinus", litval$system)] <- "pinus"
unique(litval$system[is.na(litval$main_per)])

litval$ID_AF <- NA
litval$ID_AF[grepl("orests|orest |orest$", litval$system)] <- "forest"
litval$ID_AF[grepl("onoculture", litval$system)] <- "monoculture"
litval$ID_AF[!is.na(litval$main_crop) & !is.na(litval$main_per)] <- "two-crops"
litval$ID_AF[grepl("ome garden", litval$system)] <- "HG"
unique(litval$system[is.na(litval$ID_AF)])

litval$var <- as.numeric(gsub(" ", "", tstrsplit(litval$value, "±")[[2]]))
litval$value <- gsub(" ", "", tstrsplit(litval$value, "±")[[1]])

mult_values <- which(sapply(strsplit(litval$value, ",|and"), length)>1)

new_rows = lapply(mult_values, function(i){
  vals = as.numeric(strsplit(litval$value[i], ",|and")[[1]])
  vals = vals[!is.na(vals)]
  
  x = matrix(litval[i,], ncol = ncol(litval), nrow = length(vals), byrow = TRUE)
  x[, colnames(litval) == "value"] = vals
  
  x = as.data.frame(x)
  colnames(x) = colnames(litval)
  return(x)
}
)

new_rows = do.call(rbind, new_rows) 

litval2 = rbind(new_rows, litval[-mult_values,])
litval = data.frame(system = unlist(litval2$system), 
                    pool = unlist(litval2$pool), 
                    value = unlist(litval2$value), 
                    location = unlist(litval2$location), 
                    ref = unlist(litval2$ref), 
                    main_crop = unlist(litval2$main_crop),
                    main_per = unlist(litval2$main_per), 
                    ID_AF = unlist(litval2$ID_AF), 
                    var = unlist(litval2$var))

```

# References
